---
title: 'Peer-graded Assignment: Regression Models Course Project'
author: "Alejandro Osorio"
date: "June 29, 2018"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Choosing the Model

As 

## Appendix 1: Preliminary Data Analysis

Basic properties of the 'mtcars' dataset:

```{r, echo = FALSE}
str(mtcars)
```

So, we're talking of a small sample of only 32 observations.

Additionally, the visual correlation between variables can be seen as follows:

```{r, echo = FALSE}
pairs(~ ., mtcars)
```

## Appendix 2: Choosing the Predictors

After studying the meaning of each variable, which shows it's a consecuence of the rest of variables 'qsec' (1/4 mile time) is eliminated from the analysis.

The rest of the variables make sense as potential predictors not only from their meanings, but also by the visual correlations between them and 'mpg', as observed in Appendix 1 (such as the inverse relation between 'mpg' and 'hp' or 'wt', among others). 

## Appendix 3: Choosing the Model Type

Given the nature of the analysis (continuous outcome, obtained from discrete and continuous predictors), the size of the dataset (only 32 observations), plus the visual correlations observed between 'mpg' (the outcome) and its predictors, the model to be used wil be linear (lm type).

