---
title: 'Peer-graded Assignment: Regression Models Course Project'
author: "Alejandro Osorio"
date: "June 29, 2018"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(car)
data(mtcars)
mtcars <- data.frame(mtcars)
```

## Choosing the Model

As 

## Appendix 1: Preliminary Data Analysis

Basic properties of the 'mtcars' dataset:

```{r, echo = FALSE}
str(mtcars)
```

So, we're talking of a small sample of only 32 observations.

Additionally, visual correlations among pairs of potential variables can be seen as follows:

```{r, echo = FALSE}
pairs(~ ., mtcars)
```

## Appendix 2: Model Selection

Given the nature of the analysis (continuous outcome, obtained from discrete and continuous regressors), the size of the dataset (only 32 observations), plus the visual correlations observed between 'mpg' (the outcome) and its potential regressors, the model to be used wil be linear (lm type), with 'am' as a not-so-dummy binary-factor variable.

## Appendix 3: Choosing Preliminary Regressors

After some web research (such as http://www.driverside.com/auto-library/top_10_factors_contributing_to_fuel_economy-317 and https://www.quora.com/On-what-factors-does-mileage-of-a-vehicle-depend), the main variables (and therefore candidates for main regresors), are: displacement, power, aerodynamics, weight and number of forward gears.

Besides aerodynamics (not available in the dataset), the visual correlations between them and 'mpg' are pretty clear, as seen in Appendix 1 (such as the inverse relation between 'mpg' and 'hp' or 'wt', among others). Therefore, they were the first candidates for regression modeling in the next appendix.

## Appendix 4: Regression Modeling

